Attribute fusion in a latent process model for time series of graphs

### 2011-10-30 ssp2011.tex
Draft of Journal version completed

\documentclass[final]{IEEEtran}
\usepackage[utopia]{mathdesign}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{mathrsfs}
\usepackage{bm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[noadjust]{cite}
%\usepackage[colon,sort&compress]{natbib}
%\numberwithin{equation}{section}
\renewcommand\arraystretch{1.2}
\let\underbrace\LaTeXunderbrace
\let\overbrace\LaTeXoverbrace
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathllapinternal#1#2{%
\llap{$\mathsurround=0pt#1{#2}$}% $
}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathclapinternal#1#2{%
\clap{$\mathsurround=0pt#1{#2}$}%
}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathrlapinternal#1#2{%
\rlap{$\mathsurround=0pt#1{#2}$}% $
}
\bibliographystyle{IEEEtran}
\begin{document}
\title{Attribute fusion in a latent process model for time series of
  graphs}
\author{Carey~E.~Priebe, Nam~H.~Lee, Youngser~Park, and Minh~Tang}%
%\thanks{Johns Hopkins University \\ Department of Applied Mathematics
%  and Statistics \\ Baltimore, Maryland 21218-2682 USA}%
\maketitle
\begin{abstract}
 We consider the problem of anomaly/change point detection given a
 time series of graphs with categorical attributes on the
 edges. Various attributed graph invariants are considered, and their
 power for detection as a function of a linear fusion parameter is
 presented.  
\end{abstract}
\begin{IEEEkeywords}
  Anomaly detection, Attributed Random Graphs, Fusion, Random Dot
  Product Graphs
\end{IEEEkeywords}
\section{Introduction}

@@

presented in
\cite{lee:_laten_proces_model_time_attrib_random_graph}. We summarized
here the ideas that are relevant to our discussion. We begin by
introducing some terminology. Let $\mathscr{S}$ be the unit simplex in

@@

in $[n]$ and edge attributes $\mathscr{K} = [K]
$ is then a pair $(\mathbf{X},G)$ of random elements such that

@@

\item $\mathbf{X} = \{X_i\}_{i = 1}^{n}$ is a collection of

@@

\item $G$ is a random graph with vertices set $[n]$ such that

@@

    \mathbb{P}(i \sim j \,|\, \mathbf{X} = (x_1, x_2, \dots,
    x_n)) = \langle x_i, x_j \rangle.
  and that $\mathbf{P}(i \sim j \,|\, \mathbf{X})$ and $\mathbf{P}(i' \sim
  j' \,|\, \mathbf{X})$ are independent whenever $(i,j) \not = (i',j')$. If
  $i \sim j$ in $G$, then the attribute of the edge $\{i,j\}$ is an
  element of $\mathscr{K}$. In particular, $\{i,j\}$ has attribute $k$
  with probability $x_{i,k} x_{j,k}$. 

@@

Let $\mathscr{K}_{+} = \{1,\dots,K+1\}$. We say that a c\'{a}dl\'{a}g
process $W \colon [0,\infty) \mapsto \mathscr{K}_{+}^{n}$ induces the
sequence of random dot product spaces $\mathscr{V} = \{X(t), G(t)\}_{t
  = 1}^{\infty}$ if

@@

\item Each $(X(t), G(t))$ is a random dot product space with vertices
  $[n]$ and attributes $\mathscr{K}$. Furthermore, for each
  $i \in [n]$, $k \in \mathscr{K}$ and $t \in \mathbb{N}$, we have
  $X_{i,k}(t)  = \int_{t - 1}^{t}{ \mathbf{1}\{W_i(u) = k\}\, du}$.

@@

    \mathbb{P}(G(t) = g \,|\, \mathscr{F}_{\leq t}) = \mathbb{P}(G(t) = g \,|\, X(t))
where $\mathscr{F}_{\leq t}$ are the sigma fields generated by $\{W(s)

@@

\item For each $t \in \mathbb{N}$ and vertex $i \in [n]$, there exists
  a matrix $\mathbf{Q}^{(i)}(t)$ such that $W_i$, when restricted to

@@

  chain with state space $\mathscr{K}_+$, intensity matrix
  ${\mathbf{Q}^{(i)}(t)}$, and stationary distribution
  $\pi^{(i)}(t)$. $W$ is thus a stationary, continuous-time Markov chain with state
  space $\mathscr{K}_{+}^{n}$ and intentisy matrix
  $\otimes_{i=1}^{n}\mathbf{Q}^{(i)}(t)$.
\item There exists a $t^{*} \in \mathbb{N}$ and a $m < n$ such that 

@@

  \item for  $t < t^{*}$
      \pi^{(i)}(t) \equiv \pi_0 \\
      \mathbf{Q}^{(i)}(t) \equiv \mathbf{Q}_0
  \item  for $t \geq t^{*}$

@@

change-point phenomena. We will refer to $(t^{*}, m, \pi_0, \pi_1,
\mathbf{Q}_0, \mathbf{Q}_1)$ as the change parameters. Because $W$ is completely
determined by the change parameters, we often choose to omit $W$ and
only mention $\mathscr{V}$ when referring to random dot process models
with change-point phenomena. For a rdpm $\mathscr{V}$ with change
parameters $(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$, we can construct
several approximations to $\mathscr{V}$. Of particular interests are the
following two approximations.

@@

  \bar{G}(t)\}_{t = 1}^{\infty}$ of independent random dot product

@@

 \item For $t < t^{*}$,

@@

     \bar{X}_{i}(t) \equiv \hat{\pi}_0.
 \item For $t \geq t^{*}$
     \bar{X}_{i}(t) \equiv \hat{\pi}_1 \quad \text{for $i \leq m$} \\
     \bar{X}_{i}(t) \equiv \hat{\pi}_0 \quad \text{for $i > m$} 

@@

@@

     \mathbf{Q}_1)^{-1}(\mathbf{I} - \mathbf{1}\pi_1^{T})

@@

  for the continuous-time Markov chain on $\mathscr{K}_{+}$ with intensity

@@

  e.g., \cite[p. 55]{asmussen03:_applied_probab_queues}). Let

@@

    \mathrm{diag}(\pi_1)

@@

  \item For each $t$ and each $i \in [n]$, $\widehat{X}_i(t)$ is a

@@

    vector $Z_{i}(t)$ to $\mathscr{S}$ with mean and covariance

@@

  \item For $t < t^{*}$,

@@

      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_0 \\
      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_0

@@

  \item For $t \geq t^{*}$,  
      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_0, \quad \text{for $i \leq

@@

      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_1, \quad \text{for $i > 

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{0} \quad \text{for $i

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{1} \quad \text{for $i

@@

    obtained by removing the last row and column of $\Sigma_0^{(r)}$ and
    $\Sigma_1^{(r)}$, respectively. 

@@

Suppose that we have a sequence $\{\mathscr{V}^{r} \colon r > 0 \}$ of
rdpm with vertices $[n]$ and attributes $\mathscr{K}$ where for each
$r > 0$, $\mathscr{V}^{r}$ has change parameters $(t^{*}, m, \pi_0,
\pi_1, r \mathbf{Q}_0, r \mathbf{Q}_1)$. The parameter $r$ can be
thought of as the vertex process rate, i.e., the waiting time
decreases exponentially as $r$ increases. Let us now consider the
sequence of first approximation $\bar{\mathscr{V}}^{r}$ and second
approximations $\widehat{\mathscr{V}}^{r}$ of $\mathscr{V}$. We note
that $\bar{\mathscr{V}}^{r_1} = \bar{\mathscr{V}}^{r_2}$ for any $r_1,
r_2 > 0$. Let us then denote by $\bar{\mathscr{V}}$ the (a.e.) unique
first order approximation of $\mathscr{V}^{r}$ for $r > 0$. If we
denote by $\widehat{\Sigma}_{0}^{(r)}$ and
$\widehat{\Sigma}_{1}^{(r)}$ the matrices $\widehat{\Sigma}_0$ and
$\widehat{\Sigma}_1$ for $\widehat{\mathscr{V}}^{r}$, then
$\widehat{\Sigma}_0^{(r)} = \widehat{\Sigma}_0^{(1)}/r$ and
$\widehat{\Sigma}_{1}^{(r)} = \widehat{\Sigma}_1^{(1)}/r$. Therefore,
as $r \rightarrow \infty$, $\widehat{\Sigma}_0^{(r)} \rightarrow 0$
and $\widehat{\Sigma}_1^{(r)} \rightarrow 0$ and so
$\widehat{\mathscr{V}}^{r} \rightsquigarrow
\bar{\mathscr{V}}$. This indicates that for sufficiently large $r$,
there is little statistical difference amongst the $\bar{\mathscr{V}}$,
$\mathscr{V}^{r}$ and $\widehat{\mathscr{V}}$ (\cite[Theorem
2]{lee:_laten_proces_model_time_attrib_random_graph}). $\bar{\mathscr{V}}$
will thus serve as the basis for our subsequent analysis on graph
invariants for attributed random graphs. 
\section{Change-point detection}
Let $\mathscr{V}$ be a random dot process model with change parameters
$(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. The
change-parameters encapsulate a notion of chatter anomalies, i.e., a
subset of vertices of $\mathscr{V}$ with altered communication
behaviour in an otherwise stationary setting as depicted in
Fig.~\ref{fig:notional_change_point}.
  \includegraphics[width=8cm]{graphics/Fig1-SSP2011.pdf}
  \caption{Notional depiction of the problem of change-point detection
    in a time-series of graphs}
  \label{fig:notional_change_point}
\\ \noindent
We are interested in
the problem of testing, for a $t \in \mathbb{N}$, the hypotheses that
$t$ is the change-point of $\mathscr{V}$, namely
  \mathscr{H}_0 \colon t^{*} > t \\
  \mathscr{H}_A \colon t^{*} = t
%\subsection{Graphs Invariants}
%In this paper we consider the problem of detecting chatter anomalies
This will be done using the notion of fusion of attributed graph
invariants.  The particular invariants of interests are the size
$\mathcal{E}$, number of triangles $\tau$, scan $\Psi$, and
max degree $\Delta$. Let $G = (V,E,\phi)$ be an attributed graph with 
$\phi \colon \tbinom{V}{2} \mapsto \mathscr{K}_{+}$. For a given
$\{u,v\} \in \tbinom{V}{2}$, we define a $\Gamma_{uv} \in
\mathbb{R}^{K+1}$ by $\Gamma_{uv}(k) = \mathbb{I}\{\phi(\{u,v\}) =
k\}$. Let $\widehat{\Gamma}_{uv}$ be the first $K$ coordinates of
$\Gamma_{uv}$. Thus $\widehat{\Gamma}_{uv} = \bm{0}$ unless $\{u,v\}
\in E$. Under the independent edges assumption, the
$\Gamma_{uv}$ are independent, and hence the $\widehat{\Gamma}_{uv}$
are also independent. We can consider linear attribute fusion of graph invariants with
parameter $\lambda \in \mathbb{R}^{K}$ via
%\begin{gather}
%  \label{eq:6}
%  \mathcal{E}_{\lambda}(G) = \sum_{k=1}^{K} \lambda_k \sum_{u,v}
%  \mathbb{I}\{ \phi(uv) = k \} \\
%  \tau_{\lambda}(G) = \sum_{(i,j,k) \in K^{3}\vphantom{\tbinom{V}{3}}}
%  \, \, \sum_{(u,v,w)
%      \in \tbinom{V}{3}} \lambda_{i}\lambda_j \lambda_k h_{ijk}(u,v,w) \\
%  \Delta_{\lambda}(G) = \max_{v \in V} \, \sum_{k = 1}^{K} \lambda_k
%  \sum_{u \in N(v)}{\mathbb{I}\{\phi(\{u,v\}) = k\}} \\
%  \label{eq:3}
%  \Psi_{\lambda}(G) = \max_{v \in V} \sum_{k = 1}^{K}
%  \lambda_k \!\!\sum_{u,w \in N(v)}\!\!\! \mathbb{I}\{\phi(\{u,w\}) = k\} 
%  \end{gather}
  \label{eq:6}
  \mathcal{E}_{\lambda}(G) &= \sum_{\mathclap{\{u,v\} \in
      \tbinom{V}{2}}} \langle
  \lambda, \widehat{\Gamma}_{uv} \rangle \\
  \tau_{\lambda}(G) &= \sum_{\mathclap{\{u,v,w\} \in \tbinom{V}{3}}} \langle
  \lambda, \widehat{\Gamma}_{uv} \rangle \langle \lambda,
  \widehat{\Gamma}_{uw} \rangle \langle \lambda, \widehat{\Gamma}_{vw}
  \rangle \\ 
\Delta_{\lambda}(G) &= \max_{v \in V} \, \sum_{w \in N(v)} \langle
\lambda, \widehat{\Gamma}_{vw} \rangle \\
\label{eq:3}
  \Psi_{\lambda}(G) &= \max_{v \in V} \sum_{u,w \in N(v)} \langle
  \lambda, \widehat{\Gamma}_{uw} \rangle
  Let $\{G(t)\}$ be a time series of graphs. Let $J_{\lambda}(t)$ be a
  statistic for $G(t)$ of the form as in Eq.~\eqref{eq:6} through
  Eq.~\eqref{eq:3}. We define the normalization
  $\bar{J}^{l}_\lambda(t)$ of $J_{\lambda}$ based on recent past as
  \label{eq:4}
 \bar{J}^{l}_{\lambda}(t) = \frac{1}{l}\sum_{s = 1}^{l} J_{\lambda}(t - s) 
where $l \in \mathbb{N}$ specified the width of the running-average
window. Our main interests is in the normalized fusion
statistic $T_{\lambda}^{l}(t)$ as depicted in Fig.~\ref{fig:temporal}, namely
  \label{eq:7}
 T_{\lambda}^{l}(t) = % \begin{cases}
   % J_{\lambda}(t) - J_{\lambda}(t - 1) & \text{if $m = 1$} \\
   \frac{J_{\lambda}(t) -
     \bar{J}_{\lambda}^{l}(t)}{\sqrt{\tfrac{1}{l-1}
       \sum_{s=1}^{l}(J_{\lambda}(t - s) - \bar{J}_{\lambda}^{l}(t))^2}}
%   & \text{if $m \geq 2$}
 %  \end{cases}
for $l \geq 2$. As we have mentioned earlier, we will use the first
approximation $\bar{\mathscr{V}}$ of $\mathscr{V}$ in our study of the 
asymptotic theory for these graphs invariants.   

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{fontenc}
%\usepackage{lmodern}
%\usepackage[utopia]{mathdesign}
\usepackage{graphicx}
\usepackage{xltxtra}
\setmainfont[Mapping=tex-text]{Linux Libertine}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
\ifCLASSOPTIONcompsoc
\usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
\usepackage[font=footnotesize]{subfig}
\fi
\usepackage{subfig}
\usepackage{parskip}
\usepackage{mathrsfs}
\usepackage{bm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[noadjust]{cite}
\renewcommand\arraystretch{1.2}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathllapinternal#1#2{%
\llap{$\mathsurround=0pt#1{#2}$}% $
}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathclapinternal#1#2{%
\clap{$\mathsurround=0pt#1{#2}$}%
}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathrlapinternal#1#2{%
\rlap{$\mathsurround=0pt#1{#2}$}% $
}
\bibliographystyle{IEEEtran}
\begin{document}
\title{Attribute fusion in a latent process model for time series of
  graphs}
\author{Minh~Tang, Youngser~Park, Nam~H.~Lee, and Carey~E.~Priebe%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Minh~Tang, Nam~H.~Lee
 and Carey~E.~Priebe are with the Department of Applied Mathematics
 and Statistics, Johns Hopkins University.\protect \\
  Email: \{mtang10, nhlee, cep\}@jhu.edu
\IEEEcompsocthanksitem Youngser Park is with the Center for Imaging
Science , Johns Hopkins University.\protect \\
  Email: youngser@jhu.edu}%
\thanks{}}
\markboth{IEEE Transactions on Pattern Analysis and Machine
  Intelligence (TPAMI)}{}
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
  Hypothesis testing on time series of attributed graphs has
  applications in diverse areas, e.g., social network analysis
  (wherein vertices represent individual actors or organizations),
  connectome inference (wherein vertices are neurons or brain regions)
  and text processing (wherein vertices represent authors or
  documents).  We consider the problem of anomaly/change point
  detection given the latent process model for time series of graphs
  with categorical attributes on the edges presented in \cite{lee11}.
  Various attributed graph invariants are considered, and their power
  for detection as a function of a linear fusion parameter is
  presented.  Our main result is that inferential performance in
  mathematically tractable first-order and second-order approximation
  models does provide guidance for methodological choices applicable
  to the exact (realistic but intractable) model. Furthermore, to the
  extent that the exact model is realistic, we may tentatively
  conclude that approximation model investigations have some bearing
  on real data applications.
\end{abstract}
\begin{IEEEkeywords}
  Anomaly detection, Attributed Random Graphs, Random Dot
  Product Graphs.
\end{IEEEkeywords}}
\maketitle
\section{Introduction}

@@

presented in \cite{lee11}. We summarize here the ideas that are
relevant to our discussion.  Let $G = (V,E)$ be an undirected graph,
equipped with an edge-attribution function $\phi \colon \tbinom{V}{2}
\mapsto \{1,2,\dots, K+1\}$ for some $K$ such that $\phi(e) \leq K$
for $e \in E$ and $\phi(e) \equiv K+1$ for $e \not \in E$. We refer to
$\phi(e)$ as the attribute of edge $e \in E$, and the pair $(G,\phi)$
as an attributed graph. Let $\mathscr{S}$ be the unit simplex in

@@

in $[n]$ and edge attributes $[K]$ is a pair $(\mathbf{X},G)$ of
random elements such that

@@

\item $\mathbf{X} = \{X_v\}_{v = 1}^{n}$ is a collection of

@@

\item $G$ is a random undirected graph with vertex set $[n]$ such that

@@

    \mathbb{P}(u \sim v \,|\, \mathbf{X}) = \mathbb{P}(u \sim v \, |
    \, X_u,X_v) = \langle x_u, x_v \rangle
  and $\bm{1}(u \sim v \,|\, \mathbf{X})$ and $\bm{1}(u' \sim
  v' \,|\, \mathbf{X})$ are independent whenever $(u,v) \not =
  (u',v')$. Furthermore, we also have 
    \mathbb{P}(\phi(\{u,v\} = k \, | \, u \sim
    v) = x_{u,k} x_{v,k}. 

@@

We say that a c\'{a}dl\'{a}g process $W \colon [0,\infty) \mapsto
[K+1]^{n}$ induces the sequence of random dot product spaces
$\mathscr{V} = (\mathbf{X}(t), G(t))_{t = 1}^{\infty}$ if

@@

\item Each $(\mathbf{X}(t), G(t))$ is a random dot product space with
  vertices $[n]$ and edge attributes $[K]$. Furthermore, for each $v
  \in [n]$, $k \in [K]$ and $t \in \mathbb{N}$, we have
    \label{eq:54}
  X_{v,k}(t)  = \int_{t - 1}^{t}{ \mathbf{1}\{W_v(\omega) = k\}\, d\omega}
  where we denote $W_{v}(\omega)$ as the $v$-th component of $W(\omega)$ and
  $X_{v,k}$ as the $k$-th component of $X_v$.  

@@

    \mathbb{P}(G(t) = g \,|\, \mathscr{F}_{\leq t}) = \mathbb{P}(G(t)
    = g \,|\, \mathbf{X}(t))
where $\mathscr{F}_{\leq t}$ is the $\sigma$-field generated by $\{W(s)

@@

\item For each $t \in \mathbb{N}$ and vertex $v \in [n]$, there exists
  a matrix $\mathbf{Q}^{(v)}(t)$ such that $W_v$, when restricted to

@@

  chain with state space $[K+1]$, intensity matrix
  ${\mathbf{Q}^{(v)}(t)}$, and stationary distribution
  $\pi^{(v)}(t)$. $W$ is thus a piecewise stationary, continuous-time
  Markov chain with state space $[K+1]^{n}$ and intensity matrix
  $\otimes_{v=1}^{n}\mathbf{Q}^{(v)}(t)$.
\item There exists a $t^{*} \in \mathbb{N}$ and a $m < n$ such that,
  for some $\pi_0, \pi_1, \mathbf{Q}_0$ and $\mathbf{Q}_1$ we have

@@

  \item for $t < t^{*} - 1$,
      \pi^{(1)}(t) = \dots = \pi^{(n)}(t) = \pi_0 \\
      \mathbf{Q}^{(1)}(t) = \dots = \mathbf{Q}^{(n)}(t) = \mathbf{Q}_0,
  \item  for $t \geq t^{*} - 1$,

@@

change-point phenomena at time $t = t^{*}$. We chose $\{1,2,\dots,m\}$
as the set of vertices with a change in the stationary distribution
and intensity matrix at time $t = t^{*} - 1$ for ease of
discussion. Permutation of the vertex labels does not affect our
subsequent analysis. We will refer to $(t^{*}, m, \pi_0, \pi_1,
\mathbf{Q}_0, \mathbf{Q}_1)$ as the change parameters.  Because $W$ is
completely determined by the change parameters, we often choose to
omit $W$ and only mention $\mathscr{V}$ when referring to random dot
process models (rdpm) with change-point phenomena. For a rdpm
$\mathscr{V}$ with change parameters $(t^{*}, m, \pi_0, \pi_1,
\mathbf{Q}_0, \mathbf{Q}_1)$, we can construct several approximations
to $\mathscr{V}$. Of particular interests are the following two
approximations.

@@

  \bar{G}(t)\}_{t = 1}^{\infty}$ of {\em independent} random dot product

@@

 \item for $t < t^{*} - 1$,

@@

     \bar{X}_{v}(t)  \equiv \hat{\pi}_0 \quad \text{for all $v \in [n]$}
 \item for $t \geq t^{*} - 1$
     \bar{X}_{v}(t) \equiv \hat{\pi}_1 \quad \text{for $v \leq m$} \\
     \bar{X}_{v}(t) \equiv \hat{\pi}_0 \quad \text{for $v > m$} 

@@

The first approximation yields a sequence of independent random graphs
with independent edges. For $t \leq t^{*} - 1$, $G(t)$ is an
attributed instance of the Erd\"{o}s-Renyi random graphs. For $t \geq
t^{*}$, $G(t)$ is an attributed instance of the kidney-egg model
\cite{rukhin11} (see \figurename~\ref{fig:kidney-egg}), which in
itself is a special instance of the mixed membership stochastic
blockmodel \cite{airoldi08:_mixed}. The unattributed kidney-egg model
can be denoted as $\kappa(n,m,p,q,s)$ with the quantities $p$, $q$,
and $s$ representing
p = \mathbb{P}(u \sim v \, | \, u \leq m, v \leq m) = \langle
\hat{\pi}_1, \hat{\pi}_1 \rangle \\ q =
\mathbb{P}(u \sim v \, | \, u \leq m, v > m) = \langle \hat{\pi}_1,
\hat{\pi}_0 \rangle \\ s = \mathbb{P}(u
\sim v \, | \, u > m, v > m) = \langle \hat{\pi}_0, \hat{\pi}_0
\rangle . 
\begin{figure}[!tp]
  \includegraphics[width=5cm]{graphics/Fig2-SSP2011.pdf}
  \caption{An illustration of the kidney-egg model for $G(t^{*})$.}
  \label{fig:kidney-egg}

@@

     \mathbf{Q}_1)^{-1}(\mathbf{I} - \mathbf{1}\pi_1^{T}).

@@

  for the continuous-time Markov chain on $[K+1]$ with intensity

@@

  e.g. \cite[p. 55]{asmussen03:_applied_probab_queues}). Let

@@

    \mathrm{diag}(\pi_1).

@@

  \item For each $t$ and each $v \in [n]$, $\widehat{X}_v(t)$ is a

@@

    vector $Z_{v}(t)$ to $\mathscr{S}$ with mean and covariance

@@

  \item For $t < t^{*} - 1$,

@@

      \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_0 \\
      \mathrm{Var}[Z_v(t)] \equiv \widehat{\Sigma}_0.

@@

  \item For $t \geq t^{*} - 1$,  
      \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_0, \quad \text{for $v \leq

@@

      \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_1, \quad \text{for $v > 

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{0} \quad \text{for $v

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{1} \quad \text{for $v

@@

    obtained by removing the last row and column of $\Sigma_0$ and
    $\Sigma_1$, respectively. 

@@

A second-order approximation yields a sequence of independent latent
position graphs \cite{marchette08:_predic,scheinerman10:_model,hoff02:_laten}.

Suppose that we have a sequence $\{\mathscr{V}^{r} \colon r > 0 \}$ of
rdpm with vertices $[n]$ and attributes $[K]$ where for each $r > 0$,
$\mathscr{V}^{r}$ has change parameters $(t^{*}, m, \pi_0, \pi_1, r
\mathbf{Q}_0, r \mathbf{Q}_1)$. The parameter $r$ can be thought of as
the vertex process rate, i.e., the waiting time decreases
exponentially as $r$ increases. Let us now consider the sequence of
first approximations $\bar{\mathscr{V}}^{r}$ and second approximations
$\widehat{\mathscr{V}}^{r}$ of $\mathscr{V}^{r}$. We note that
$\bar{\mathscr{V}}^{r_1} = \bar{\mathscr{V}}^{r_2}$ for any $r_1, r_2
> 0$. Let us then denote by $\bar{\mathscr{V}}$ the (a.e.) unique
first order approximation of $\mathscr{V}^{r}$ for $r > 0$. If we
denote by $\widehat{\Sigma}_{0}^{(r)}$ and
$\widehat{\Sigma}_{1}^{(r)}$ the matrices $\widehat{\Sigma}_0$ and
$\widehat{\Sigma}_1$ for $\widehat{\mathscr{V}}^{r}$, then
$\widehat{\Sigma}_0^{(r)} = \widehat{\Sigma}_0^{(1)}/r$ and
$\widehat{\Sigma}_{1}^{(r)} = \widehat{\Sigma}_1^{(1)}/r$. Therefore,
as $r \rightarrow \infty$, $\widehat{\Sigma}_0^{(r)} \rightarrow
\bm{0}$ and $\widehat{\Sigma}_1^{(r)} \rightarrow \bm{0}$ and so
$\widehat{\mathscr{V}}^{r} \overset{\mathrm{d}}{\longrightarrow}
\bar{\mathscr{V}}$. This indicates that for sufficiently large $r$,
there is little statistical difference amongst the
$\bar{\mathscr{V}}$, $\mathscr{V}^{r}$ and $\widehat{\mathscr{V}}^{r}$
(\cite[Theorem 2]{lee11}). The behavior of
$\widehat{\mathscr{V}}^{r}$ for ``small'' $r$ is less
clear. $\bar{\mathscr{V}}$ will thus serve as the basis for our
subsequent analysis on graph invariants for attributed random graphs.
\subsection{Change-point detection}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=8cm]{graphics/Fig1-SSP2011.pdf}
  \caption{Notional depiction of the problem of change-point detection
    in a time-series of graphs}
  \label{fig:notional_change_point}
\end{figure}
Let $\mathscr{V}$ be a random dot process model with change parameters
$(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. The
change-parameters encapsulate a notion of chatter anomalies, i.e., a
subset of vertices of $\mathscr{V}$ with altered communication
behavior in an otherwise stationary setting as depicted in
Fig.~\ref{fig:notional_change_point}. We are interested in the problem
of testing, for a $t \in \mathbb{N}$, the hypotheses that $t$ is the
change-point of $\mathscr{V}$, namely
\begin{gather*}
  \mathscr{H}_0 \colon t^{*} > t \\
  \mathscr{H}_A \colon t^{*} = t
\end{gather*}
This will be done using the notion of fusion of attributed graph
invariants. The particular invariants of interests are the size
$\mathcal{E}$, number of triangles $\tau$, scan $\Psi$, and max degree
$\Delta$. The scan statistics $\Psi$ are introduced and applied to the
problem of detecting chatter anomalies in
\cite{priebe05:_scan_statis_enron_graph}. The use of various graph
invariants as test statistics was considered in
\cite{pao11:_statis_infer_random_graph} and it was shown there, via
Monte Carlo analysis, that no single invariant is uniformly most
powerful.

Let $(G,\phi)$ be an attributed graph. For a given $\{u,v\}
\in \tbinom{V}{2}$, we define a $\Gamma_{uv} \in \mathbb{R}^{K}$ by
$\Gamma_{uv}(k) = \mathbb{I}\{\phi(\{u,v\}) = k\}$. Thus $\Gamma_{uv}
= \bm{0}$ unless $\{u,v\} \in E$. Under the independent edges
assumption, the $\Gamma_{uv}$ are independent. We consider linear
attribute fusion of graph invariants with parameter $\lambda \in
\mathbb{R}^{K}$ via
%\begin{gather}
%  \label{eq:6}
%  \mathcal{E}_{\lambda}(G) = \sum_{k=1}^{K} \lambda_k \sum_{u,v}
%  \mathbb{I}\{ \phi(uv) = k \} \\
%  \tau_{\lambda}(G) = \sum_{(i,j,k) \in K^{3}\vphantom{\tbinom{V}{3}}}
%  \, \, \sum_{(u,v,w)
%      \in \tbinom{V}{3}} \lambda_{i}\lambda_j \lambda_k h_{ijk}(u,v,w) \\
%  \Delta_{\lambda}(G) = \max_{v \in V} \, \sum_{k = 1}^{K} \lambda_k
%  \sum_{u \in N(v)}{\mathbb{I}\{\phi(\{u,v\}) = k\}} \\
%  \label{eq:3}
%  \Psi_{\lambda}(G) = \max_{v \in V} \sum_{k = 1}^{K}
%  \lambda_k \!\!\sum_{u,w \in N(v)}\!\!\! \mathbb{I}\{\phi(\{u,w\}) = k\} 
%  \end{gather}

\begin{align}
  \label{eq:6}
  \mathcal{E}_{\lambda}(G) &= \sum_{\mathclap{\{u,v\} \in
      \tbinom{V}{2}}} \langle
  \lambda, \Gamma_{uv} \rangle \\
  \tau_{\lambda}(G) &= \sum_{\mathclap{\{u,v,w\} \in \tbinom{V}{3}}} \langle
  \lambda, \Gamma_{uv} \rangle \langle \lambda,
  \Gamma_{uw} \rangle \langle \lambda, \Gamma_{vw}
  \rangle \\ 
\Delta_{\lambda}(G) &= \max_{v \in V} \, \sum_{w \in N(v)} \langle
\lambda, \Gamma_{vw} \rangle \\
\label{eq:3}
  \Psi_{\lambda}(G) &= \max_{v \in V} \sum_{u,w \in N[v]} \langle
  \lambda, \Gamma_{uw} \rangle.
  \end{align}
  where $N(v) = \{u \colon u \sim v\}$ is the set of neighbors of
  $v$ and $N[v] = v \cup N(v)$.

  Let $\{G(t)\}$ be a time series of graphs. Let $J_{\lambda}(t)$ be a
  statistic for $G(t)$ of the form as in Eq.~\eqref{eq:6} through
  Eq.~\eqref{eq:3}. We define the running average 
  $\bar{J}^{l}_\lambda(t)$ of $J_{\lambda}$ as
\begin{equation}
  \label{eq:4}
 \bar{J}^{l}_{\lambda}(t) = \frac{1}{l}\sum_{s = 1}^{l} J_{\lambda}(t - s) 
\end{equation}
where $l \in \mathbb{N}$ specified the width of the running-average
window. Our main interest is in the normalized fusion statistic
$T_{\lambda}^{l}(t)$ as depicted in Fig.~\ref{fig:temporal}, namely
\begin{equation}
  \label{eq:7}
 T_{\lambda}^{l}(t) = % \begin{cases}
   % J_{\lambda}(t) - J_{\lambda}(t - 1) & \text{if $m = 1$} \\
   \frac{J_{\lambda}(t) -
     \bar{J}_{\lambda}^{l}(t)}{\sqrt{\tfrac{1}{l-1}
       \sum_{s=1}^{l}(J_{\lambda}(t - s) - \bar{J}_{\lambda}^{l}(t))^2}}
%   & \text{if $m \geq 2$}
 %  \end{cases}
\end{equation}
for $l \geq 2$, i.e., we want to find a parameter $\lambda$ such that
the power of the test using $T_{\lambda}^{l}$ is maximum. We note that
$T_{\lambda}^{l}$ is scale invariant in $\lambda$ for all of our graph
invariants. It was noted in \cite{lee11} that the maximum of the power
for scale invariant tests is attainable in $\{ \lambda \colon \|
\lambda \| = 1 \}$. \\ \\

As we have mentioned earlier, we will assume that the number of
vertices $n$ is large and that the use of the first order
approximation $\bar{\mathscr{V}}$ of $\mathscr{V}$ is appropriate in
our study of the asymptotic theory for these graphs invariants. We
strive to obtain approximations for the power in testing $t^{*} > t$
against $t^{*} = t$ using $T_{\lambda}(t)$ and from these
approximations, find the $\lambda$ that maximize the power. We now
introduce additional notation that will be used later on in the
paper. First, let $\pi_{00}, \pi_{01}$, and $\pi_{11}$ be
sub-probability vectors whose components are given by
